Calculating Biological Module Enrichment or Depletion and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2016-01-13 jss/article.tex
[skip ci] summary

>>>>

\section[Discussion]{Discussion}
In this work, we presented the R package \pkg{ACSNMineR}, a novel package for
the calculation of p-values for enrichment or depletion of genes in biological
pathways (also called modules here). The main function of the package,
\code{enrichment}, offers many options to set the statistical test used (either
the Fisher exact test or the hypergeometric test), the modules (or maps) used,
the universe of reference for the calculations, the miminal size for a module to
be taken into account, the choice for over or under-representation. Additional
functions allow users to create graphical representations for p-values
(\code{represent\_enrichment}) or to perform calculations for multiple samples
or groups (\code{multisample\_enrichment}). Large-scale molecular maps of the
Atlas of Cancer Signaling Networks (ACSN) are included in the package (six
maps), but any other molecular pathway (such as those defined in the REACTOME
database) can be imported and used for calculations with the command
\code{format\_from\_gmt}.

We also describe in this work the \pkg{RNaviCell} package, a R package highly
complementary to \pkg{ACSNMineR}. This package is dedicated to create data
visualization on ACSN maps, and is communicating with the NaviCell web server to
create interactive graphical representations. Users might use this tools to
represent genes of interest that have been shown to be related to the maps by
calculating enrichment with the  \pkg{ACSNMineR}.  Creating maps with the
graphical user interface of the ACSN website can be a tedious task if the user
has multiple samples or gene lists, and wants to compare their representations
on ACSN maps. The \pkg{RNaviCell} package can be used to automate the process of
creating the graphical representations automatically.

Of course, \pkg{ACSNMineR} is not the only R package for enrichment calculations. For
instance, \pkg{GOstats} \citep{falcon2007using} is probably one of the first
packages that was created to calculate enirchment for Gene Ontology categories.
\pkg{GOstats} can also be used to calculate enrichment for other biological
pathways categories, such as KEGG pathways (by using an instance of the class
\code{KEGGHyperGParams}) or PFAM protein families (using
\code{PFAMHyperGParams}). However, its usage might not be as straightforward as
\pkg{ACSNMineR}, and it does not seems possible to test user-defined biological
pathways. Furthermore, other authors have pointed out that the KEGG database
used by this package has not been updated since 2012. \pkg{clusterProfiler} is a
recent R package released for the
